---
title: '20220922: tidy data'
author: "wyy"
date: "2022-09-27"
output: github_document
---
```{r}
library(tidyverse)
options(tibble.print_min = 5)
```
#   pivot_longer
In data import, we used the haven package to load the PULSE biomarkers dataset from a .sas7bdat. Let’s reload those data and take a closer look:
```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()

pulse_df
```
With our new understanding of tidy data, we quickly recognize a problem: the BDI score is spread across four columns, which correspond to four observation times. We can fix this problem using pivot_longer:
```{r}
pulse_tidy_data = 
  pivot_longer(
    pulse_df, 
    bdi_score_bl:bdi_score_12m, #column
    names_to = "visit", #column name
    values_to = "bdi")#value name

pulse_tidy_data
```
This looks much better! However, now visit is an issue. The original column names were informative but we probably don’t need to keep the bdi_score_ prefix in each case. I’ll use an additional option in pivot_longer to address this:

```{r}
pulse_tidy_data = 
  pivot_longer(
    pulse_df, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi")

pulse_tidy_data
```
In the preceding I’ve saved intermediate datasets to make each step clear. While this can be a helpful crutch as you’re trying out code, it is generally bad practice. There are also some additional transformations needed to wrap up the data wrangling process, like changing bl to 00m for consistency across visits and converting visit to a factor variable. (It’s possible that you would want visit to be a numeric variable instead, which could be done with a different call to mutate.) Lastly, it’s nice to organize the data into a reasonable order.

Altogether, then, the code below will import, tidy, and transform the PULSE dataset into a usable format:
```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  relocate(visit) %>%
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit)) %>%
  arrange(id, visit)

print(pulse_df, n = 12)
```
#   Learning Assessment: 
In the litters data, the variables gd0_weight and gd18_weight give the weight of the mother mouse on gestational days 0 and 18. Write a data cleaning chain that retains only litter_number and these columns; produces new variables gd and weight; and makes gd a numeric variable taking values 0 and 18 (for the last part, you might want to use recode …). Is this version “tidy”?
```{r}
litter_df = 
  read_csv("./data/FAS_litters.csv") %>%
  janitor::clean_names()%>%
  select(litter_number, ends_with("weight")) %>% 
  pivot_longer(
    gd0_weight:gd18_weight,
    names_to = "gd",
    values_to = "weight"
  )%>%
  mutate(gd = recode(gd, "gd0_weight" = 0, "gd18_weight" = 18))#recode:gd0_weight全部改成0

litter_df
```
#   pivot_wider
We’ve been exclusively interested in tidying data, but we’ve admitted that sometimes untidy is better for human consumption. For that reason we’re going to take a short digression into untidying your tidy data.

The code below creates a tidy dataset that could result from an analysis. This is the correct format for additional analysis or visualization, but doesn’t facilitate quick comparisons for human readers.

```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)

analysis_result
```

```{r}
pivot_wider(
  analysis_result, 
  names_from = "time", #column name
  values_from = "mean")#value
```
We’re pretty much there now – in some cases you might use select to reorder columns, and (depending on your goal) use knitr::kable() to produce a nicer table for reading.

#   Binding rows
We’ve looked at single-table non-tidy data, but non-tidiness often stems from relevant data spread across multiple tables. In the simplest case, these tables are basically the same and can be stacked to produce a tidy dataset. That’s the setting in LotR_words.xlsx, where the word counts for different races and genders in each movie in the trilogy are spread across distinct data rectangles (these data are based on this example).

To produce the desired tidy dataset, we first need to read each table and do some cleaning.

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>%
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")
```
Here it was necessary to add a variable to each dataframe indicating the movie; that information had stored elsewhere in the original spreadsheet. As an aside, the three code snippets above are all basically the same except for the range and the movie name – later we’ll see a better way to handle cases like this by writing our own functions, but this works for now.

Once each table is ready to go, we can stack them up using bind_rows and tidy the result:
```{r}
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) %>% 
  #bind_rows: 3 tables have the same column
  janitor::clean_names() %>%
  pivot_longer(
    female:male,
    names_to = "gender", 
    values_to = "words") %>%
  mutate(race = str_to_lower(race)) %>% #race转小写
  select(movie, everything()) 

lotr_tidy
```
#   Joining datasets
```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv") %>%
  janitor::clean_names() %>%
  mutate(
    sex = recode(sex, `1` = "male", `2` = "female"),
    sex = factor(sex)) 
```

```{r}
litter_data = 
  read_csv("./data/FAS_litters.csv") %>%
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  #sep=3: separate after 3 characters
  relocate(litter_number) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    dose = str_to_lower(dose))
```

```{r}
fas_data = 
  left_join(pup_data, litter_data, by = "litter_number")

fas_data
```
#   anti_join
anti_join() return all rows from x without a match in y.

#   Learning Assessment: 
The datasets in this zip file contain de-identified responses to surveys included in past years of this course. Both contain a unique student identifier; the first has responses to a question about operating systems, and the second has responses to questions about degree program and git experience. Write a code chunk that imports and cleans both datasets, and then joins them.